---
title: "ESM206 Final"
author: "Teresa Fukuda"
date: "12/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages and data}

# load packages

library(tidyverse) # wrangle and look at data 
library(janitor) # clean up names to snakecase yay
library(car) # ANOVA
library(effsize) # compute effect sizes
library(corrplot) # For correlation matrices

# load data

ca_county_pesticides <- read_csv("ca_county_pesticides.csv")
climate_opinion <- read_csv("climate_opinion.csv")
pesticide_residues<- read_csv("PesticideResidues.csv")
tmas_df <- read_csv("tmas_df.csv")
ucsb_climate <- read_csv("ucsb_climate.csv")


```


### Task 1. Climate Change Opinions

```{r Task 1}

# For Task 1: Explore relationships between climate change perspectives, information exposure, education attainment, poverty levels and coastline on the proportion of people who believe that global warming is happening (i.e., ‘happening’ should be your dependent variable).

# Exploratory data visualization

climate_opinion_vis <- ggplot (climate_opinion)+
  geom_point(aes(y=happening, x=poverty_rate)) +
  geom_point(aes(y=happening, x=bachelors), color="red") +
  geom_point(aes(y=happening, x=mediaweekly), color="blue") +
  geom_point(aes(y=happening, x=discuss), color="orange") +
  geom_point(aes(y=happening, x=consensus), color="green")
  
climate_opinion_vis

# create a df with only the numeric data
pair_df <- climate_opinion[2:7]

# look at the pairwise graphs of the data from pairs_df
pairs(pair_df)
# could be linear relationship for most; the poverty_rate plots look more blob than linear

# create a correlation matrix df to use in the corrplot
cor_df <- cor(pair_df)

# visually look at this using corrplot function to make correlation matrix
corrplot(cor_df, #input matrix of correlation data, not original data
         type= "upper",# only shows upper half of matrix (so as not to repeat data)
         method= "ellipse",
         tl.col="black", # change label text color and size
         tl.cex=0.5) 

```

```{r linear model}



```


a. A brief statement (1-sentence) explaining the type of analysis you chose to perform, and why it is appropriate.

b. An exploratory correlation matrix for all continuous variables (make it readable, but it does not need to be finalized and you do not need to add a figure caption).
c. A finalized table of results for your analyses (you may present more than one model), with a figure caption.

d. Choose two coefficients in the model that you find interesting (either because they are expected or unexpected). For each, write a single sentence explaining what that coefficient means in terms of an expected change in the dependent variable.

e. Based on diagnostic plots, describe (1 - 2 sentences) if you have concerns about heteroscedasticity or normality of residuals for your model(s). You do not need to include the diagnostic plots in your final document.

```{r}

```

